Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeastern Taiwan

### 2020-01-03 analysis/paper/paper.Rmd
add sentences to introduction

## social changes in cross-cultural interaction contexts
- here I have the third cultural resistance, identity take a look at social boundary
- hypothesis, goals
- trade [@Alizadeh2018]

Culture contact between indigenous people and colonizer with imperial power usually leads to profound changes in local indigenous societies. We explore the culture changes in indigenous societies in northeastern Taiwan that experienced foreign interactions with the Europeans and Chinese since 17th century by examining the local pottery production. We observed ceramic production in northeastern Taiwan in the late Iron Age period and historical period (1400-1900 AD) presenting a high consistency in form and shape compared to other pottery found from other regions on this island that hints the emergence of pottery specialization. Our aim is to explore the impact of foreign contacts on local indigenous societies in northeastern Taiwan by examining pottery production as an indicator for social changes. Using pottery shape as a proxy to study craft specialization, we answer these questions: Did foreign contacts have impacts on indigenous pottery production that can be detected in the shape of the vessels? How does pottery change in shape throughout 600 years at this site? Did pottery become more homogeneous and standardized in shape after foreign contacts with the European colonizer or the Chinese immigrant?

@@

## Measuring ceramic standardization

@@

- identity, social boundary

>>>>

The emergence of social inequality in the context of cross-cultural interaction especially in a colonial situation was observed in many parts of world when the introduction of foreign trade goods to local indigenous societies. The monopolization for long-distance trade goods might cause substantial transformations of indigenous economic, cultural, and socio-political systems [@Dietler2005; @Dietler1997; @Junker1993; @Silliman2005]. The archaeological evidence in Northeastern Taiwan show some extent of social differentiation when encountered by the Spanish and the Dutch in the 17th century based on the distribution pattern of the trade ornaments that were viewed as prestige goods in local historical context. In addition to the differential distribution of trade ornaments in domestic area and burial contexts, the locally made ceramic present a high consistency in form and shape compared to other pottery found from other regions on this island that hints the possibility of pottery specialization [@Chen2007]. To further understand the relationship among the use of prestige goods, the degree of ceramic specialization and the arrival of foreign colonizers, this paper uses pottery production as an example for identifying social changes in a cross-cultural contact context.

<---[@Acabado2018] In order to maintain status, it was important for elites to monopolize foreign trade and control the manufacture of earthenware pots connected to it through centralized and specialized production. --->

@@

<--Culture contact between indigenous people and colonizer with imperial power usually leads to profound changes in local indigenous societies. We explore the culture changes in indigenous societies in northeastern Taiwan that experienced foreign interactions with the Europeans and Chinese since 17th century by examining the local pottery production. We observed ceramic production in northeastern Taiwan in the late Iron Age period and historical period (1400-1900 AD) presenting a high consistency in form and shape compared to other pottery found from other regions on this island that hints the emergence of pottery specialization. Our aim is to explore the impact of foreign contacts on local indigenous societies in northeastern Taiwan by examining pottery production as an indicator for social changes. Using pottery shape as a proxy to study craft specialization, we answer these questions: Did foreign contacts have impacts on indigenous pottery production that can be detected in the shape of the vessels? How does pottery change in shape throughout 600 years at this site? Did pottery become more homogeneous and standardized in shape after foreign contacts with the European colonizer or the Chinese immigrant? -->

## Measuring ceramic standardization
- intentional and mechanical

@@

## social changes in cross-cultural interaction contexts
- cultural resistance
- trade [@Alizadeh2018]
- identity, social boundary, Raux2015, Boness2017
